---
title: "ADA Final Project"
author: "Byeongju Ryu"
date: "2024-12-01"
output: html_document
---

## README: Introduction
I will use the 2016 Health and Retirement Study to answer two questions:
  1. Is there an association between residential relocation and loneliness among community-dwelling adults aged 50 and over in the U.S.?
  2. Does volunteer status moderate the assocation above?

I used the cleaned dataset for the analyses. The HRS datasets are publicly available on the official HRS data (https://hrsdata.isr.umich.edu/). I used Stata/MP 18.5 to clean the dataset. If you have questions about data cleaning process, please contact: ryubyeongju@wustl.edu

The variable names in the dataset are consistent with the HRS codebook (https://www.rand.org/well-being/social-and-behavioral-policy/portfolios/aging-longevity/dataprod/hrs-data.html).

## Set the working directory
```{r}
setwd("/Users/ryu/Library/CloudStorage/Box-Box/HRS Projects/Relocation, volunteering, loneliness/data and dofile")
```

## Load the dataset
```{r}
# Load the library
library(haven)

# Load the dataset: We have 4,109 observations and 16 variables in this cleaned dataset.
data <- read_dta ("HRS_cleaned.dta")

# Display the first few rows of the dataset
head(data)
```

## Look at the descriptive statistics
```{r}
# Load the libraries
library(tableone)

# List of categorical variables
categorical_vars <- c(
  "r13lblonely3_quintile_dummy4",    ## Loneliness (1 = Yes, 0 = No)
  "r13relocation_region",            ## Relocation status (1 = Yes, 0 = No)
  "r13fvolB",                        ## Volunteer status (1 = Yes, 0 = No)
  "female",                          ## Gender (1 = Female, 0 = Male)
  "race_binary",                     ## Race (1 = White, 0 = Non-White)
  "r13married",                      ## Marital Status (1 = Married/Partnered, 0 = No)
  "r13ivolB",                        ## Informally helping friends/neighbors/relatives (1 = Yes, 0 = No)
  "r13urbrurB"                       ## Rurality (1 = Rural, 0 = Urban/Suburban)
)

# List of continuous variables
continuous_vars <- c(
  "raedyrs",                         ## Education (years)
  "r13agey_b",                       ## Age (years)
  "h13itot",                         ## Total household income (annual, $)
  "r13adl5a",                        ## Number of ADL limitations (0-5)
  "r13closechild_topcoded"           ## Number of close children (0-5)
)

# Combine all variables
all_vars <- c(categorical_vars, continuous_vars)

# Create the TableOne object
summary_table <- CreateTableOne(vars = all_vars, data = data, factorVars = categorical_vars)

# Print the table
print(summary_table)
```

## Apply the HRS survey weights
```{r}
# Load survey package
library(survey)

# Define the survey design
survey_design <- svydesign(
  ids = ~hhid,              # Specify PSU (primary sampling unit: household)
  weights = ~r13wtresp,     # Specify sampling weights
  data = data               # Specify the dataset
)

# Print the survey design object to check
summary(survey_design)
#
```

## Define survey design with subpopulation
```{r}
# Define the subpopulation (where 'included' indicates the subpopulation)
subpop_design <- subset(survey_design, included == 1)
```

## Run survey-weighted logistic regression (Research Question 1)
```{r}
# Convert variables to factors for categorical variables
data$r13relocation_region <- factor(data$r13relocation_region)
data$r13fvolB <- factor(data$r13fvolB)
data$female <- factor(data$female)
data$race_binary <- factor(data$race_binary)
data$r13married <- factor(data$r13married)
data$r13ivolB <- factor(data$r13ivolB)
data$r13urbrurB <- factor(data$r13urbrurB)
data$r13adl5a <- factor(data$r13adl5a)

# Fit the logistic regression model
logistic_model <- svyglm(
  r13lblonely3_quintile_dummy4 ~ 
    r13relocation_region + 
    r13fvolB + 
    r13agey_b + 
    female + 
    race_binary + 
    raedyrs + 
    r13married + 
    h13itot + 
    r13adl5a + 
    r13closechild_topcoded + 
    r13ivolB + 
    r13urbrurB,
  design = subpop_design,
  family = quasibinomial()
)

# Display the summary of the logistic regression model
summary(logistic_model)

# Extract coefficients, confidence intervals, and p-values
coefficients <- coef(logistic_model)
conf_intervals <- confint(logistic_model)
p_values <- summary(logistic_model)$coefficients[, "Pr(>|t|)"]

# Combine into a single data frame
results <- data.frame(
  Variable = names(coefficients),
  OddsRatio = exp(coefficients),
  LowerCI = exp(conf_intervals[, 1]),
  UpperCI = exp(conf_intervals[, 2]),
  PValue = p_values
)

# Display results
results
```

## Calculate the predicted probabilities (Research Question 1)
```{r}
# Load libraries
library(margins)

# Compute margins for r13relocation_region (relocation status)
margins_region <- predicted_data %>%
  group_by(r13relocation_region) %>%
  summarise(
    Margin = mean(predicted_prob, na.rm = TRUE),   # Average predicted probability
    StdErr = sd(predicted_prob, na.rm = TRUE) / sqrt(n()), # Standard error
    LowerCI = Margin - 1.96 * StdErr,             # Lower 95% confidence interval
    UpperCI = Margin + 1.96 * StdErr              # Upper 95% confidence interval
  )

# Display margins for r13relocation_region (relocation status)
print(margins_region)

# Compute margins for r13fvolB (volunteer status)
margins_volunteer <- predicted_data %>%
  group_by(r13fvolB) %>%
  summarise(
    Margin = mean(predicted_prob, na.rm = TRUE),   # Average predicted probability
    StdErr = sd(predicted_prob, na.rm = TRUE) / sqrt(n()), # Standard error
    LowerCI = Margin - 1.96 * StdErr,             # Lower 95% confidence interval
    UpperCI = Margin + 1.96 * StdErr              # Upper 95% confidence interval
  )

# Display margins for r13fvolB (volunteer status)
print(margins_volunteer)
```

## Run survey-weighted logistic regression (Research Question 2)
```{r}
# Fit the logistic regression model with interaction terms
logistic_model <- svyglm(
  r13lblonely3_quintile_dummy4 ~ 
    r13relocation_region * r13fvolB +  # Interaction term
    r13agey_b + 
    female + 
    race_binary + 
    raedyrs + 
    r13married + 
    h13itot + 
    r13adl5a + 
    r13closechild_topcoded + 
    r13ivolB + 
    r13urbrurB,
  design = subpop_design,
  family = quasibinomial()
)

# Display the model summary
summary(logistic_model)

# Extract coefficients, confidence intervals, and p-values
coefficients <- coef(logistic_model)
conf_intervals <- confint(logistic_model)
p_values <- summary(logistic_model)$coefficients[, "Pr(>|t|)"]

# Combine results into a data frame
results <- data.frame(
  Variable = names(coefficients),
  OddsRatio = exp(coefficients),  # Exponentiate coefficients to get odds ratios
  LowerCI = exp(conf_intervals[, 1]),  # Lower bound of confidence interval
  UpperCI = exp(conf_intervals[, 2]),  # Upper bound of confidence interval
  PValue = p_values  # P-values
)

# Display the results
results

```

## Calculate the predicted probabilities (Research Question 2)
```{r}
# Generate predicted probabilities
predicted_data <- data %>%
  mutate(predicted_prob = predict(logistic_model, type = "response"))

# Compute margins for the interaction term
margins_interaction <- predicted_data %>%
  group_by(r13relocation_region, r13fvolB) %>%
  summarise(
    Margin = mean(predicted_prob, na.rm = TRUE),
    StdErr = sd(predicted_prob, na.rm = TRUE) / sqrt(n()),
    LowerCI = Margin - 1.96 * StdErr,
    UpperCI = Margin + 1.96 * StdErr
  )

# Display interaction margins
print(margins_interaction)
```